---
title: "Oh canada, Birds"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE,fig.width = 12,fig.height = 9)
```


```{r load data an packages}
bird_counts <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-18/bird_counts.csv")

library(tidyverse)
library(gganimate)
library(ggthemr)

ggthemr('flat')
```


# Year with How many counted, How many counted by Hour and Total Hours

```{r Year vs How many counted with species label}
p1<-bird_counts %>%
    ggplot(.,aes(year,how_many_counted,color=how_many_counted_by_hour,
                 label=species))+
          geom_point()+labs(color="Counted by Hour")+
          xlab("Year")+ylab("How Many Counted")+
          ggtitle("How many Counted vs Year")

plotly::ggplotly(p1)
```


```{r Year vs Total Hours with species label}
p2<-bird_counts %>%
    ggplot(.,aes(year,total_hours,color=how_many_counted_by_hour,
                 label=species))+
          geom_point()+labs(color="Counted by Hour")+
          xlab("Year")+ylab("Total Hours")+
          ggtitle("Total Hours vs Year")

plotly::ggplotly(p2)
```

## Year and Total Hours by decade

```{r Year vs Total Hours}
bird_counts_new<-bird_counts
bird_counts_new$year<-cut(bird_counts_new$year,
                          breaks=c(1920,1929,1939,1949,
                                   1959,1969,1979,1989,
                                   1999,2009,2017),
                          labels=c("1920s","1930s","1940s","1950s",
                                   "1960s","1970s","1980s","1990s",
                                   "2000s","2010s"))

  bird_counts_new %>%
  replace_na(list(total_hours = 0)) %>%
  group_by(year) %>%
  summarise(mean_hours=mean(total_hours),sum_hours=sum(total_hours)) %>%
  ggplot(.,aes(year,sum_hours))+geom_col()+
         xlab("Year")+ylab("Summation of Total Hours \n(Mean Total Hours)")+
         ggtitle("Total Hours by Decade")+
         scale_y_continuous(expand=c(0,25000))+
         geom_text(aes(label=sum_hours),size=3.5,vjust=-0.5)+
         geom_text(aes(label=round(mean_hours,2)),vjust=1)

```


## Year and How many Counted by decade

```{r Year vs How many Counted}
bird_counts_new<-bird_counts
bird_counts_new$year<-cut(bird_counts_new$year,
                          breaks=c(1920,1929,1939,1949,
                                   1959,1969,1979,1989,
                                   1999,2009,2017),
                          labels=c("1920s","1930s","1940s","1950s",
                                   "1960s","1970s","1980s","1990s",
                                   "2000s","2010s"))

  bird_counts_new %>%
  replace_na(list(how_many_counted = 0)) %>%
  group_by(year) %>%
  summarise(mean_counted=mean(how_many_counted),
            sum_counted=sum(how_many_counted)) %>%
  ggplot(.,aes(year,sum_counted))+geom_col()+
         xlab("Year")+ylab("Summation of How Many Counted \n(Mean of How many counted)")+
         ggtitle("Total of How many counted")+
         scale_y_continuous(expand=c(0,75000))+
         geom_text(aes(label=sum_counted),size=3.5,vjust=-0.5)+
         geom_text(aes(label=round(mean_counted,2)),vjust=1)

```

## Year and How many Hours Counted by Hour for decade


```{r Year vs How many Counted by hour}
bird_counts_new<-bird_counts
bird_counts_new$year<-cut(bird_counts_new$year,
                          breaks=c(1920,1929,1939,1949,
                                   1959,1969,1979,1989,
                                   1999,2009,2017),
                          labels=c("1920s","1930s","1940s","1950s",
                                   "1960s","1970s","1980s","1990s",
                                   "2000s","2010s"))

  bird_counts_new %>%
  replace_na(list(how_many_counted_by_hour = 0)) %>%
  group_by(year) %>%
  summarise(mean_counted_hour=mean(how_many_counted_by_hour),
            sum_counted_hour=sum(how_many_counted_by_hour)) %>%
  ggplot(.,aes(year,sum_counted_hour))+geom_col()+
         xlab("Year")+ylab("Summation of How Many Counted by hour \n(Mean of How many counted by hour)")+
         ggtitle("Total of How many counted by hour")+
         scale_y_continuous(expand=c(0,750))+
         geom_text(aes(label=round(sum_counted_hour,2)),size=3.5,vjust=-0.5)+
         geom_text(aes(label=round(mean_counted_hour,2)),vjust=1)

```

# Top 10 Species with how many counted 

```{r Top 10 Species with how many counted}
  bird_counts_new %>%
  replace_na(list(how_many_counted = 0)) %>%
  group_by(species) %>%
  summarise(sum_counted=sum(how_many_counted)) %>%
  top_n(10) %>%
  ggplot(.,aes(str_wrap(species,12),sum_counted))+geom_col()+
         xlab("Year")+ylab("Summation of How Many Counted")+
         ggtitle("Total of How many counted")+
         geom_text(aes(label=sum_counted),size=3.5,vjust=0.5)
```


## Top 10 Species with how many counted by year

```{r Top 10 species with how many counted by year}
p1<-bird_counts%>%
    subset(species=="American Crow" | species=="Canada Goose" | 
           species=="Common Goldeneye" | species=="Common Merganser" |
           species=="European Starling" | species=="Greater Scaup" | 
           species=="Herring Gull" | species=="House Sparrow" |
           species=="Long-tailed Duck" | species=="Mallard") %>%
  replace_na(list(how_many_counted = 0)) %>%
  group_by(year,species) %>%
  ggplot(.,aes(str_wrap(species,12),how_many_counted))+geom_col()+
         xlab("Year")+ylab("How Many Counted")+
         ggtitle("How many counted for Top 10 Species",
                 subtitle ="Year:{round(frame_time)}" )+
         transition_time(year)+ease_aes("linear")+
         geom_text(aes(label=how_many_counted),size=3.5,vjust=0.5)

animate(p1,nframes=94,fps=1)
```

## Top 10 Species with how many counted by decade

```{r Top 10 species with how many counted by decade}
bird_counts_new%>%
    subset(species=="American Crow" | species=="Canada Goose" | 
           species=="Common Goldeneye" | species=="Common Merganser" |
           species=="European Starling" | species=="Greater Scaup" | 
           species=="Herring Gull" | species=="House Sparrow" |
           species=="Long-tailed Duck" | species=="Mallard") %>%
  replace_na(list(how_many_counted = 0)) %>%
  group_by(year,species) %>%
  summarise(hmc=sum(how_many_counted)) %>%
  ggplot(.,aes(str_wrap(species,12),hmc))+geom_col()+
         xlab("Year")+ylab("How Many Counted")+
         ggtitle("How many counted for Top 10 Species",
                 subtitle ="Decade:{closest_state}" )+
         transition_states(year)+ease_aes("linear")+
         geom_text(aes(label=hmc),size=3.5,vjust=0.5)
```


# Top 10 How many counted by hours

```{r Top 10 Species with how many counted by hours}
  bird_counts_new %>%
  replace_na(list(how_many_counted_by_hour = 0)) %>%
  group_by(species) %>%
  summarise(sum_counted=sum(how_many_counted_by_hour)) %>%
  top_n(10) %>%
  ggplot(.,aes(str_wrap(species,12),sum_counted))+geom_col()+
         xlab("Year")+ylab("Summation of How Many Counted by Hour")+
         ggtitle("Total of How many counted by Hour")+
         geom_text(aes(label=round(sum_counted,2)),size=3.5,vjust=0.5)
```


## Top 10 Species with how many counted by hours for year

```{r Top 10 species with how many counted by hours for year}
p1<-bird_counts%>%
    subset(species=="American Crow" | species=="Canada Goose" | 
           species=="Common Goldeneye" | species=="Common Merganser" |
           species=="European Starling" | species=="Greater Scaup" | 
           species=="Herring Gull" | species=="House Sparrow" |
           species=="Long-tailed Duck" | species=="Mallard") %>%
  replace_na(list(how_many_counted_by_hour = 0)) %>%
  group_by(year,species) %>%
  ggplot(.,aes(str_wrap(species,12),how_many_counted_by_hour))+geom_col()+
         xlab("Year")+ylab("How Many Counted by Hour")+
         ggtitle("How many counted by Hour for Top 10 Species",
                 subtitle ="Year:{round(frame_time)}" )+
         transition_time(year)+ease_aes("linear")+
         geom_text(aes(label=round(how_many_counted_by_hour,2)),size=3.5,vjust=0.5)

animate(p1,nframes=94,fps=1)
```

## Top 10 Species with how many counted by hours for decade

```{r Top 10 species with how many counted by hours for decade}
bird_counts_new%>%
    subset(species=="American Crow" | species=="Canada Goose" | 
           species=="Common Goldeneye" | species=="Common Merganser" |
           species=="European Starling" | species=="Greater Scaup" | 
           species=="Herring Gull" | species=="House Sparrow" |
           species=="Long-tailed Duck" | species=="Mallard") %>%
  replace_na(list(how_many_counted_by_hour = 0)) %>%
  group_by(year,species) %>%
  summarise(hmc=sum(how_many_counted)) %>%
  ggplot(.,aes(str_wrap(species,12),hmc))+geom_col()+
         xlab("Year")+ylab("How Many Counted")+
         ggtitle("How many counted for Top 10 Species",
                 subtitle ="Decade:{closest_state}" )+
         transition_states(year)+ease_aes("linear")+
         geom_text(aes(label=round(hmc,2)),size=3.5,vjust=0.5)
```


*THANK YOU* 

