---
title: "week 35 bridges of baltimore"
author: "M.amalan"
date: "November 27, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.width = 9,fig.height = 7)

library(readr)
library(tidyverse)
library(stringr)
library(ggthemr)
library(gganimate)
library(formattable)

ggthemr("flat dark")

bridges <- read_csv("baltimore_bridges.csv")
#View(bridges)

names(bridges)<-c("lat","long","County","Carries","Year Built","Condition","Average Daily Traffic","Total Improvement","Month","Year","Owner","Responsibility","Vehicles")
attach(bridges)
```

# Bridge Data and Baltimore 

Data on bridges is week 35 for TidyTuesday. Trying to explain the data using is obvious, yet I
will use animated jitter plots. There are 13 variables and 2079 observations. Brave choice of 
limiting my self to less than 10 variables, where latitude, longitude and Vehicles will not be
taken into account.

So with the help of packages tidyverse, ggthemr, gganimate,formattable and readr I will complete this
analysis. Most of the bridges are owned by several agencies, but I will only focus on the 
top three ownership holders.

## Counties which have bridges owned by State Highway Agency 

Close to 1000 bridges are owned by State Highway Agency, where most of them are in Baltimore
County. High amount of bridges are in good condition, further more bridges are in Fair condition and
only around 10 bridges in Poor condition. 

Considering the Average Daily Traffic only one bridge in Poor condition has the amount of 
close to 110,000, while all the other poor condition bridges have Average Daily Traffic less
than 30,000. While counties Anne Arundel and Hartford have no Poor condition bridges at all.

Most of the bridges are from Baltimore County and around 20 bridges have more than 150,000
Average Daily Traffic for both Fair and Good conditions. Hartford and Carroll Counties have 
their Average Daily Traffic which does not exceed 80,000 at any condition of the bridge.

```{r County with Condition and Average Daily Traffic SHA}
ggplot(subset(bridges,Owner=="State Highway Agency")
       ,aes(color=Condition,y=`Average Daily Traffic`,x=str_wrap(County,7)))+xlab("County")+
  ggtitle("Condition of Bridges owned by State Highway Agency \nand their Average Daily Traffic")+
  scale_y_continuous(labels =seq(0,230000,10000) ,breaks = seq(0,230000,10000))+
  geom_jitter()+transition_states(Condition,transition_length = 2,state_length = 3)+
  enter_fade()+exit_shrink()+ease_aes("back-in")
```

## Counties which have bridges owned by  County Highway Agency 

County Highway Agency owns the second  most amount of bridges in this data-set. 
Therefore using we are going to check how the condition of the bridge 
and counties are explained is the simplest manner. 

Less amount of poor condition bridges in all counties except Anne Arundel
County. All bridges owned by County Highway Agency have a limited Average 
Daily Traffic less than 50,000. Clearly we have More fair bridges than good 
ones. In the poor condition category only two have Average Daily Traffic
more than 20,000, while other two have more than 10 bridges.

Most of these bridges are in Baltimore County even it is in any one of three
conditions. There are few bridges which have more than 40,000 Average Daily
Traffic and they are also in Baltimore County.

There are bridges which have Zero Average Daily Traffic. In all three Conditions
only Hartford County has bridges which has Average Daily Traffic less than 
10,000.

```{r County with Condition and Average Daily Traffic CHA}
ggplot(subset(bridges,Owner=="County Highway Agency")
       ,aes(color=Condition,y=`Average Daily Traffic`,x=str_wrap(County,7)))+xlab("County")+
  ggtitle("Condition of Bridges owned by County Highway Agency \nand their Average Daily Traffic")+
  scale_y_continuous(labels =seq(0,40000,5000) ,breaks = seq(0,40000,5000))+
  geom_jitter()+transition_states(Condition,transition_length = 2,state_length = 3)+
  enter_fade()+exit_shrink()+ease_aes("back-in")
```

##  Counties which have bridges owned by  State Toll Authority

There is only bridge which is in Poor condition and it is in Baltimore County, while counties
Howard and Anne Arundel have no Good condition bridges. Further there is only 3 Fair condition
bridges in Howard County while there have an Average Daily Traffic less than 10,000. 

The highest Average Daily Traffic is close to 170,000 which are only 4 and in Good and Fair conditions.
Further, Anne Arundel County has only one Good bridge and is Hartford it is six bridges. Only few 
of bridges have Average Daily Traffic close to zero. 

```{r County with Condition and Average Daily Traffic STA}
ggplot(subset(bridges,Owner=="State Toll Authority")
       ,aes(color=Condition,y=`Average Daily Traffic`,x=str_wrap(County,7)))+xlab("County")+
  ggtitle("Condition of Bridges owned by State Toll Authority \nand their Average Daily Traffic")+
  scale_y_continuous(labels =seq(0,170000,10000) ,breaks = seq(0,170000,10000))+
  geom_jitter()+transition_states(Condition,transition_length = 2,state_length = 3)+
  enter_fade()+exit_shrink()+ease_aes("back-in")
```

## Most amount of bridges Built based on Year 

Years 1957, 1970, 1975, 1991, 1963 and 1961 have the top 6 spots of for building more than 50 bridges
in those years. If we consider the conditions of Fair and Good only the year 1991 is suitable to mention,
while all other years has at-least one Poor condition bridge. Further There are more Poor condition bridges 
in 1961 than in 1957. While all Poor condition bridges has Average Daily Traffic less than 50,000. 

Finally, there are only a few bridges which have Average Daily Traffic above 100,000 and only 3 are in Good
condition. There are Bridges which can have Average Daily Traffic close to zero in all 6 years. 

```{r Year Built with Condition and Average Daily Traffic}
ggplot(subset(bridges,`Year Built`=="1957" | `Year Built`=="1970" | `Year Built`=="1975" | `Year Built`=="1991" 
              | `Year Built`=="1963" | `Year Built`=="1961")
       ,aes(color=Condition,y=`Average Daily Traffic`,x=factor(`Year Built`)))+
  xlab("Year Built")+ylab("Average Daily Traffic")+
  ggtitle("Most amount of Bridges built based on Years \nand their Conditions")+
  geom_jitter()+legend_bottom()+transition_states(Condition,transition_length = 2,state_length = 3)+
  enter_fade()+exit_shrink()+ease_aes("back-in")
```

## Average Traffic Less than or equal to 100,000 for Counties with Bridge Condition

While obtaining summary for county variable there is one issue because there are two observations
which say "Baltimore city" than "Baltimore City" and I don't want to change them. 

If we focus of Average Daily Traffic less than or equal to 100,000 based on County and Condition. 
It is clear that Poor condition bridges are part of this criteria and mostly Average Daily Traffic 
is less than 5000 for Counties Howard, Hartford and Carroll. While Baltimore County has highest amount 
up-to 75,000, but Baltimore County has highest amount close to 40,000. Finally Anne Arundel County has only
one Poor condition bridge which has Average Daily Traffic Close to zero. 

We can see that there are more Fair Condition bridges than Good ones. In Baltimore County most of the Fair
condition bridges have Average Daily Traffic less than 15000. Similarly Carroll county and Hartford county 
also behave similarly. But for Good condition bridges this is not the case where there is no certain strong
dense region as similar to Fair condition bridges. 

Previously when we looked into Ownership we did not see Baltimore City as a factor, but here that is not the
case. 

```{r Average Traffic less than 100000}
ggplot(subset(bridges, `Average Daily Traffic` <= 100000 & County != "Baltimore city"),aes(x=County,y=`Average Daily Traffic`,color=Condition))+
  xlab("County")+ylab("Averag Daily Traffic")+ggtitle("Average Daily Traffic Less than 100,000 \nFor Counties")+
  scale_y_continuous(labels = seq(0,100000,5000),breaks = seq(0,100000,5000))+coord_flip()+
  theme(axis.text.x = element_text(angle = -90))+
  geom_jitter()+transition_states(Condition,transition_length = 2,state_length = 3)+
  enter_fade()+exit_shrink()+ease_aes("back-in")

```

## Average Traffic More than  100,000 for Counties with Bridge Condition

This Jitter plot is completely different than previous one, because there are no clear dense regions 
for any counties and conditions of the bridge. There is only one Poor condition bridge in Baltimore County
where the Average Daily Traffic is close to 115,000. In Fair condition bridges also Baltimore County holds 
the most, while they are slightly dense in the region of 175,000 to 190,000. while for Howard County similar
density occurs between 190,000 to 205,000. Bridges in Good condition have a more center in Baltimore County
than Anne Arundel County because there are no bridges in that criteria. 

```{r  Average Traffic more than 100000}
ggplot(subset(bridges, `Average Daily Traffic` > 100000 & County != "Baltimore city"),aes(x=County,y=`Average Daily Traffic`,color=Condition))+
  xlab("County")+ylab("Average Daily Traffic")+ggtitle("Average Daily Traffic More than 100,000 \nFor Counties")+
  scale_y_continuous(labels = seq(100000,230000,5000),breaks = seq(100000,230000,5000))+coord_flip()+
  theme(axis.text.x = element_text(angle = -90))+
  geom_jitter()+transition_states(Condition,transition_length = 2,state_length = 3)+
  enter_fade()+exit_shrink()+ease_aes("back-in")

```


## Improvement and Bridge Conditions with Counties 

In the variable of Total Improvement There are 1438 missing values, 42 values are zero and 
the rest are actual values. I am going to look at Total Improvement in two tables. First 
table will include where bridges have Total Improvement is higher than 9999,000 and less than 
30,000,000. Second table is for bridges which have Total Improvement higher than or equal to 
30,000,000. 

Further to make these tables interesting I will using the package formattable package and 
all colors and tiles for numerical values. In the first table there are 7 bridges while 
only Anne Arundel County holds 3 and Baltimore City holds 4. One bridge is from 1953, and others
are from the period of 1977 to 1983. Conditions of these bridges re mostly Fair and two bridges 
are in Good condition. Lowest Average Daily Traffic is 11760, while highest is 124193, where both bridges 
are in Fair Condition, and the amount spent on them for Total Improvement are respectively 18163
and 16264. The bridge with Highest amount of traffic is built in 1953.


```{r Bridge Condition and County with Improvement Next 7}
Top10<-subset(bridges[,c(-1,-2,-9,-10,-11,-12,-13)], `Total Improvement` > 9999 & `Total Improvement` < 30000)

customRed0 = "#FF8080"
customRed = "#7F0000"

customyellow0 = "#FFFF80"
customyellow = "#BFBF00"

customblue0 = "#6060BF"
customblue =  "#00007F"

formattable(Top10,align=c("l","l","c","c","c","c"),
            list(
              County =formatter("span",style= ~style(color="grey")),
            `Total Improvement`=color_tile(customblue0,customblue),
            `Average Daily Traffic`=color_tile(customyellow0,customyellow),
            `Year Built`=color_tile(customRed0,customRed)
            ))
```

When I did try to plot the top ten bridges with most Total improvement there 
was one issue, which is the distance between first two values and the next 8 values.
Therefore I divided the table into to. 

In this second table We can see there are two bridges which are from Baltimore City
and are built in 1980 and 1971, but the amount spent on Total Improvement is 
300,000,000. But there Average Daily Traffic is respectively 56280 and 30600.

While we have another bridge from Baltimore City and built in 1907, but Total
Improvement amount is 35,026,000. But the Average Daily Traffic is 3,900,000.


```{r Bridge Condition and County with Improvement Top 3}
Top3<-subset(bridges[,c(-1,-2,-9,-10,-11,-12,-13)], `Total Improvement` >= 30000)

customRed0 = "#FF8080"
customRed = "#7F0000"

customyellow0 = "#FFFF80"
customyellow = "#BFBF00"

customblue0 = "#6060BF"
customblue =  "#00007F"

formattable(Top3,align=c("l","l","c","c","c","c"),
            list(
              County =formatter("span",style= ~style(color="black")),
            `Total Improvement`=color_tile(customblue0,customblue),
            `Average Daily Traffic`=color_tile(customyellow0,customyellow),
            `Year Built`=color_tile(customRed0,customRed)
            ))
```
