---
title: "Malaria deaths"
author: "M.amalan"
date: "November 13, 2018"
output: 
        html_document :
          github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggrepel)
library(ggthemr)
ggthemr("flat")
load("Malaria.RData")
```

Packages : ggplot2, ggrepel, ggthemr
TidyTuesday : week 33
Data: malaria_deaths

Malaria_deaths_plot: Plot shows the death rate per 100,000 people in sri lanka decreasing rapidly from 1996 to 2003 with a drop of
                     0.89 to 0.29, and by 2016 it reaches 0. While in 2015 this rate is 0.13. Highest death rate was in 1990 with 
                     0.91.

```{r Deaths}
#data subset has been used
#scales of x and y have been more scrutinized
#labels have been added
#x axis have been modified to accomodate the years 
Malaria_deaths_plot<-ggplot(subset.data.frame(malaria_deaths,Code=="LKA"),
                      aes(x=Year,y= `Deaths - Malaria - Sex: Both - Age: Age-standardized (Rate) (per 100,000 people)`,
                      label=round(`Deaths - Malaria - Sex: Both - Age: Age-standardized (Rate) (per 100,000 people)`,2)))+
                      geom_point()+geom_line()+geom_text_repel()+
                      ggtitle("Malaria Deaths for Both genders in a rate over the years in Sri Lanka")+
                      ylab("Deaths - Malaria - Sex: Both -\n Age: Age-standardized (Rate) (per 100,000 people)")+
                      scale_y_continuous(breaks=seq(0.1,1,by=0.1) ,labels=seq(0.1,1,by=0.1))+
                      scale_x_continuous(breaks=seq(1990,2016),labels =seq(1990,2016))+
                      theme(axis.text.x = element_text(angle = 90))

print(Malaria_deaths_plot)

ggsave(Malaria_deaths_plot,width = 10,height = 10,dpi=300,filename = "Malaria_Deaths_Sri Lanka.png")
```

Data:  malaria_deaths_age
Malaria_deaths_age_plot: There are five categories in concern,where age category 15-49 has the most counts of in the range of
                          45-50. Second category is Under 5 close to 35 counts, while third category is 50-69 inbetween 20-25. 
                          It should be noted that this  order is for the year 1990. At the end of year 2015 this is not the case,
                          where the categories and counts are 15-49 (close to 10), 50-69 (less than 10), 70 or order(close to 5),
                          under 5(less than 5) and finally 5-14 (close to 0).
                          

```{r Deaths age}
#data subset has been used
#according to age group colors are assigned
#scales of x and y have been more scrutinized
#labels have been added
#x axis have been modified to accomodate the years 
Malaria_deaths_age_plot<-ggplot(subset.data.frame(malaria_deaths_age,code=="LKA"),
                                aes(x=year,y=deaths,color=factor(age_group)))+
                         geom_point()+geom_line()+
                         ggtitle("Malaria Deaths by age category in Sri Lanka over the years")+
                         ylab("Deaths Count")+
                         scale_x_continuous(breaks=seq(1990,2016),labels =seq(1990,2016))+
                         scale_y_continuous(breaks=seq(0,60,by=5) ,labels=seq(0,60,by=5))+
                         theme(axis.text.x = element_text(angle = 90))+
                         scale_color_discrete(name="Age Category")

print(Malaria_deaths_age_plot)

ggsave(Malaria_deaths_age_plot,width=10,height = 10,dpi = 300,filename = "Malaria Deaths Sri Lanka by Age.png")
```
