---
title: "Week 8 Phds In USA"
author: "M.Amalan"
date: "February 19, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 12,fig.height = 12,warning = FALSE,message = FALSE)
```


```{r load the packages and data}
# load the packages
library(tidyverse)
library(ggthemr)
library(readr)
library(gganimate)
library(ggridges)
library(ggalluvial)

ggthemr("flat")

#load the data
phdlist <- read_csv("phd_by_field.csv")
```

Five variables are representing this entire data-set and three of them are factors while 
one column represents the year and the final column is for counts. There are few missing values.
We can focus on Phds awarded from 2008 to 2017 in perspective of Broad Field, Major Field and 
Field.

[Dataset](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-02-19)

Broad Field and Major Field are considered specially but not the column Field as 
it would be difficult to plot based on the amount of categories.

# Broad Field 

Broad field has 7 categories and clearly Psychology and social sciences has produced
more than 4000 Phds each and every year. Which is twice comparing to other categories.
If we drop Psychology and Social sciences, now the changes over the years for other
categories are clear. 

There are more outliers in the field of Life sciences where some programs produce 
more than 1000 Phds each year comparatively to the rest categories. Except Life
sciences other categories tend to behave rarely as above producing more than 1000 Phds.

Engineering field has the lowest distribution with relative to other categories according 
to the box plot in every year.

## All fields

```{r Broad field boxplot}
p<-ggplot(phdlist,aes(x=str_wrap(broad_field,20),y=n_phds))+
          geom_boxplot()+
          xlab("Broad Field")+ylab("No of Phds")+
          transition_time(year)+ease_aes("linear")+
          ggtitle("Boxplot to Number of Phds in Broad Field",
              subtitle = "Year : {round(frame_time)}")+
          theme(axis.text.x = element_text(hjust=1,angle = 90))

animate(p,nframes=9,fps=1)
```

## Dropping Psychology and Social Sciences

```{r Broad field boxplot without psy and soc sciences}
p<-ggplot(subset(phdlist,broad_field != "Psychology and social sciences"),
          aes(x=str_wrap(broad_field,20),y=n_phds))+
          geom_boxplot()+
          xlab("Broad Field")+ylab("No of Phds")+
          transition_time(year)+ease_aes("linear")+
          ggtitle("Boxplot to Number of Phds in Broad Field",
              subtitle = "Year : {round(frame_time)}")+
          theme(axis.text.x = element_text(hjust=1,angle = 90))

animate(p,nframes=9,fps=1)

```


# Major Field

Focus now solely switch towards the Major Field column and here also we can see the strong 
outlier in Psychology over the years. Further, Physics and Astronomy field also has a very
strong outlier where over the years it reaches 2000 Phds.

Without dropping Psychology  we can see the odd behavior from the fields "Education Research",
"Economics" and "Computer and Information Sciences". Specially the gradual decrease of 
"Education Research" from 2008 to 2017.

Also in "Computer and Information Sciences" field there is an odd increase in 2012. 

After dropping the "Psychology" field we can now clearly see how other Major fields behave
over the years.

## Major Fields with Box plot

```{r major field boxplot}
p<-ggplot(phdlist,aes(x=str_wrap(major_field,20),y=n_phds,fill=broad_field))+
          geom_boxplot()+coord_flip()+
          xlab("Major Field")+ylab("No of Phds")+
          transition_time(year)+ease_aes("linear")+
          ggtitle("Boxplot to Number of Phds in Major Field",
              subtitle = "Year : {round(frame_time)}")+
          theme(axis.text.x = element_text(hjust=1,angle = 90),
                legend.position = "bottom")+
          labs(fill="Broad Field")

animate(p,nframes=9,fps=1)
```

## Major Fields without Psychology

```{r Major field without psy}
q<-ggplot(subset(phdlist,major_field != "Psychology"),
          aes(x=str_wrap(major_field,20),y=n_phds,fill=broad_field))+
          geom_boxplot()+coord_flip()+
          xlab("Major Field")+ylab("No of Phds")+
          transition_time(year)+ease_aes("linear")+
          ggtitle("Boxplot to Number of Phds in Major Field",
              subtitle = "Year : {round(frame_time)}")+
          theme(axis.text.x = element_text(hjust=1,angle = 90),
                legend.position = "bottom")+
          labs(fill="Broad Field")

animate(q,nframes=9,fps=1)
```


# Mathematics and Computer Sciences

I am a Statistics student with a glimpse of Computer science background, therefore my
next intention is to focus on the Broad Field "Mathematics and Compute Sciences". 

## Mathematics and Computer Science as a Broad field

Mathematics and Statistics has a gradual increase until 2012, but wavers higher and lower 
in the next years, but in 2016 there is a sudden increase of which would lead to around 700 
Phds awarded. Next year this decreases to 500 Phds. 

Comparing the 2 major fields "Computer and Information Sciences" with "Mathematics and Statistics"
indicate the strong gap between them awarding Phds. "Computer and Information Sciences" award 
more than twice the amount of Phds what "Mathematics and Statistics" award each year.

"Computer and Information Sciences" also hold a clear pattern with the Phds awarded.

```{r mathematics and cs bar chart}
subset(phdlist,broad_field == "Mathematics and computer sciences") %>%
      
ggplot(.,aes(x=factor(year),y=n_phds,fill=major_field))+
       geom_bar(stat="identity",position = "dodge")+
       theme(legend.position = "bottom")+
       xlab("Major Field")+ylab("Number of Phds")+
       ggtitle("Number of Phds awarded under Mathematics and CS",
               subtitle = "Year : 2008 to 2017")+
      scale_y_continuous(breaks=seq(0,1700,100),labels=seq(0,1700,100))+
          labs(fill="Major Field")
       
```


## Major Field for Mathematics and Computer Science

Box plot indicates the clear variation among these two major fields over years which 
could be used for comparison. The sudden peak in year 2012 for "Computer and Information Sciences"
interests me alot. It should be noted that "Mathematics and Statistics" has more outliers 
than "Computer and Information Sciences".

```{r major field boxplot with maths and cs}
p<-ggplot(subset(phdlist,broad_field == "Mathematics and computer sciences"),
          aes(x=str_wrap(major_field,20),y=n_phds))+
          geom_boxplot()+
          xlab("Major Field")+ylab("No of Phds")+
          transition_time(year)+ease_aes("linear")+
          ggtitle("Boxplot to Number of Phds in Major Field",
              subtitle = "Year : {round(frame_time)}")

animate(p,nframes=9,fps=1)
```

Below is a ridge plot to describe the same thing which would clearly indicate the data spread.

```{r ridge major fields}
ggplot(subset(phdlist,broad_field == "Mathematics and computer sciences"),
          aes(y=str_wrap(major_field,20),x=n_phds))+
          geom_density_ridges()+
          xlab("No of Phds")+ ylab("Major Field")+
          theme(legend.position = "bottom")+
          ggtitle("Ridge plot for Major Fields in Mathematics and Computer Sciences",
                  subtitle = "Year : 2008 to 2017")
```

##  Fields for Major Field Mathematics and Computer Science

considering the sub categories of the chosen broad field in a box plot did not
work quite well, but Computer Science phds being awarded with highest amount would
indicate the boom of Artificial Intelligence.

```{r field boxplot with maths and cs}
p<-ggplot(subset(phdlist,broad_field == "Mathematics and computer sciences"),
          aes(x=str_wrap(field,20),y=n_phds,fill=major_field))+
          geom_boxplot()+coord_flip()+
          xlab("Field")+ylab("No of Phds")+
          transition_time(year)+ease_aes("linear")+
          ggtitle("Boxplot to Number of Phds in Field",
              subtitle = "Year : {round(frame_time)}")+
          theme(legend.position = "bottom")+
          labs(fill="Major Field")

animate(p,nframes=9,fps=1)

```

To get a clear view here is the ridge plot, where Computer Science is very strong for 
"Computer and Information Sciences". It should be noted though there is only three other 
fields in this major field which are "Information Science systems", 
"Computer and Information Sciences, other" and "Computer and Information sciences, general".

More than 10 fields for the Major field "Mathematics and Statistics", where higher counts 
occur to "Statistics(Mathematics)", "Applied mathematics" and "Mathematics and Statistics,general".
Still non of these fields have passed the 1000 Phds awarded mark.

```{r ridge plot fields}
ggplot(subset(phdlist,broad_field == "Mathematics and computer sciences"),
          aes(y=str_wrap(field,20),x=n_phds,fill=major_field))+
          geom_density_ridges()+
          xlab("No of Phds")+ ylab("Field")+
          theme(legend.position = "bottom")+
          ggtitle("Ridge plot for Fields in Mathematics and Computer Sciences",
                  subtitle = "Year : 2008 to 2017")+
          labs(fill="Major Field")
```

# Major Field, Field For Mathematics and Computer Sciences

Finally an alluvial diagram just to look at the impact of Computer science field with 
respective to each year and major field.

```{r broad and major and field}
data.frame(subset(phdlist,broad_field=="Mathematics and computer sciences")) %>%
           na.omit() %>%
ggplot(.,aes(axis2=factor(str_wrap(year,10)), axis1= factor(str_wrap(major_field,10)), 
             axis3= factor(field), y=as.numeric(n_phds)))+
       scale_x_discrete(limits=c("Major Field","Year","Field"),expand = c(.05, .05))+
       geom_alluvium(aes(fill=factor(major_field)),width = 1/2)+
       geom_stratum(width=1/2,fill="white",color="grey")+
       geom_text(stat = "stratum", label.strata = TRUE)+
       theme(legend.position = "bottom")+ylab("No of Phds")+
       ggtitle("Major Field and Fields For Years 2008 to 2017",
               subtitle="Mathematics and Computer Science")+
          labs(fill="Major Field")
```


*THANK YOU*