---
title: "Medium_week_36"
author: "M.amalan"
date: "December 4, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE,fig.height = 10,fig.width = 10 )
```

```{r Preparing to TidyTuesday}
#loading packages
#load data
library(readr)
#manipulate data
library(dplyr)
library(magrittr)
# format table with expense
library(formattable)
# knitting the document
library(knitr)
# another type of table
library(kableExtra)
# playing with strings
library(stringr)
# combining two plots
library(grid)
library(gridExtra)
# that theme you wanted
library(ggthemr)
# text analysis
library(tm)
library(SnowballC)
library(RColorBrewer)
library(wordcloud)

# adding theme called fresh for plots
ggthemr("fresh")
#loading the data
medium <- read_csv("medium_datasci.csv")
attach(medium)
```

Focusing on Claps with Authors and publications, where does writing alot of posts
will get you popularity and claps. The code will focus on Top 10 Authors with 
most of the posts and Claps they have received. Further, does having an image in the 
post matter ?. Finally, word clouds for Top 10 authors and Top 5 publications with their titles.

# Claps 

Table indicates that 25,729 posts have 0 claps, while 7,093  posts with only one 
clap, and finally 3044 posts with 2 claps. The only odd one is posts with 50 claps
where the count is 970.

```{r Most Claps table top 15}
# extracting the top 15 with most claps
claps_table<-table(claps) %>%
                     sort() %>%
                     tail(15)

# table it up 
#kable(t(claps_table),"html") %>%
#  kable_styling(bootstrap_options = "striped",full_width = T) %>%
#  row_spec(0,bold = T,font_size = 13,color = "grey")
```

# Top 10 Authors and Claps for their posts

There are only two posts which do not have Images in their content. The highest 
number of claps is 60,000, a post written by Sophia Ciocca under the title 
"How Does Spotify Know You So Well?". Second Place is for the article "Blockchain
is not only crappy technology but a bad vision for the future" which was written 
by Kai Stinchombe with 53,000 claps. De Xun is the only author who has two articles 
which are in this list on the places 8 and 9 with claps respectively 37,000 and 
36,000. 

```{r Claps for Authors,titles,subtitles and Image}
# seperate medium with author, titles, claps and image
claps_A_I<-medium[,c("author","title","subtitle","claps","image")] %>%
              arrange(claps) %>%
              tail(10)
names(claps_A_I)<-c("Author","Title","Subtitle","Claps","Image")

# table it
formattable(claps_A_I[,-3],align=c("l","l","r","c"),
            list(
              Claps=color_tile("lightblue","blue"),
              Image=color_tile("red","green")
            ))
```

# Top 10 Authors with most posts

Yves Mulkers has most amount of posts with 487 but only 3,779 claps. This is not the 
case for Corsairs publishing where for 156 posts the number of claps are 111,501. 
Even looking at other authors names we can see that writing alot of posts does not create
claps.

```{r Top 10 Authors in a table}
# finding who are the top 10 authors with claps
# summary.factor(medium$author) %>%
#  sort() %>%
#  tail(11)

# extracting posts only from the top 10 authors with most posts
Top10_author<-subset(medium,
       author =="C Gavilanes" | author == "Jae Duk Seo" |
       author == "Corsair's Publishing" | author=="Alibaba Cloud" |
       author == "Ilexa Yardley" | author == "Peter Marshall" |
       author == "AI Hawk" | author == "DEEP AERO DRONES" |
       author == "Synced" | author == "Yves Mulkers")

# plotting the top 10 authors
g1_Top10_a<-ggplot(Top10_author,aes(x=author))+
            coord_flip()+ geom_bar()+
            xlab("Author")+ylab("Frequency")+
            ggtitle("Top 10 Authors and number of posts they have written")+
            geom_text(stat = 'count',aes(label=..count..),hjust=0.5)

# plotting the top 10 authors and their claps
g2_Top10_a<-Top10_author[,c("author","claps")] %>%
            group_by(author) %>%
            summarise_each(funs(sum)) %>%
            ggplot(.,aes(x=author,y=claps))+ geom_bar(stat = "identity")+
            ggtitle("Total number of Claps they got")+
            xlab("Author")+ylab("Claps")+
            coord_flip()+geom_text(aes(label=claps),hjust =0.5 )

# plotting two plots at same grid
grid.arrange(g1_Top10_a,g2_Top10_a,ncol=1)
```

# Top 10 Authors who have posts with Images

Extracting the top 10 authors with posts which have images it clear
most of the posts do have Images and they do generate claps. This is true 
for Corsairs's Publishing. With 154 posts it generates 109,906 claps. There 
are authors who have written more posts than totally received claps . It should 
be noted that two Authors did not add any Images for their post and 
they are Peter Marshall and C Gavilanes.

```{r Top 10 Authors when there is Image}
# plotting top 10 authors with Image
I1_g1_Top10_a<-subset(Top10_author[,c("author","image")],image==1) %>%
            ggplot(.,aes(x=author))+ geom_bar()+coord_flip()+
            xlab("Author")+ylab("Frequency")+  
            ggtitle("Top 10 Authors and posts which has Images")+
            geom_text(stat='count',aes(label=..count..),hjust =0.5 )

# plotting the claps for top 10 authors with Image 
I1_g2_Top10_a<-subset(Top10_author[,c("author","claps","image")],image==1) %>%
               group_by(author,image) %>%
               summarise_each(funs(sum)) %>%
               ggplot(.,aes(x=author,y=claps))+ geom_bar(stat = "identity")+
               coord_flip()+
               xlab("Author")+ylab("Claps")+
               ggtitle( "Total number of Claps they got")+
               geom_text(aes(label=claps),hjust =0.5 )    

# top plots one grid
grid.arrange(I1_g1_Top10_a,I1_g2_Top10_a,ncol=1)
```

# Top 10 Authors who have posts without Images

Posts without images have very low amount of total claps. To be specific 14 posts 
by Jae Duk Seo have 1833 claps but 2 posts by Corsair's publishing has 1595 claps. 
That is very Impressive. Further there are even posts which have claps less than 10 
where the number of posts is less than 5. 

```{r Top 10 Authors when there is No Image}    
# plotting top 10 authors with No Image
I0_g1_Top10_a<-subset(Top10_author[,c("author","image")],image==0) %>%
               ggplot(.,aes(x=author))+ geom_bar()+coord_flip()+
               xlab("Author")+ylab("Frequency")+  
               ggtitle("Top 10 Authors and posts without Images")+
               geom_text(stat='count',aes(label=..count..),hjust =0.5 )

# plotting the claps for top 10 authors with No Image 
I0_g2_Top10_a<-subset(Top10_author[,c("author","claps","image")],image==0) %>%
               group_by(author,image) %>%
               summarise_each(funs(sum)) %>%
               ggplot(.,aes(x=author,y=claps))+ geom_bar(stat = "identity")+
               coord_flip()+
               xlab("Author")+ylab("Claps")+               
               ggtitle("Total number of claps they got")+
               geom_text(aes(label=claps),hjust =0.5 )

# top plots one grid
grid.arrange(I0_g1_Top10_a,I0_g2_Top10_a,ncol=1)
  
```

# Top 10 Authors and Reading time for their posts

Reading in minutes, does it has anything to do with number of posts?. Looking at 
the plot it is clear that posts from Synced has more total reading time than Yves Mulkers
with highest number posts. The difference between posts is close 150, while difference 
between reading times is above 150 for Yves Mulkers and Synced. 

```{r Top 10 Authors and reading time for their posts}
# plotting top 10 authors with reading times
RT_g1_Top10_a<-Top10_author[,c("author","reading_time")] %>%
            group_by(author) %>%
            summarise_each(funs(sum)) %>%
            ggplot(.,aes(x=author,y=reading_time))+ geom_bar(stat = "identity")+
            ggtitle("Reading Time")+
            xlab("Author")+ylab("Reading Time in minutes")+  
            coord_flip()+geom_text(aes(label=reading_time),hjust =0.5 )

# printing the above plot with number of posts 
grid.arrange(g1_Top10_a,RT_g1_Top10_a,ncol=1)

```

# Top 5 Publications with most posts

Publications with most number of posts has the highest number of claps and order achieved  
in the "Top 5 publications and number of posts they have written" plot is maintained in the 
"Total number of Claps they got" plot as well. This simply refers, when you write alot of posts 
under a publication you will receive alot of claps because of the foundation this specific 
publications holds in Medium.

```{r Top 5 publications in a table}
# finding who are the top 5 publications with claps
# summary.factor(medium$publication) %>%
#  sort() %>%
#  tail(11)

# extracting posts only from the top 5 publications with most posts
Top5_pub<-subset(medium,
       publication =="Towards Data Science" | 
       publication == "Hacker Noon" |
       publication == "Becoming Human: Artificial Intelligence Magazine" |
       publication =="Chatbots Life" |
       publication == "Data Driven Investor" )

# plotting the top 5 publications
g1_Top5_p<-ggplot(Top5_pub,aes(x=str_wrap(publication,15)))+
            coord_flip()+ geom_bar()+
            xlab("Publication")+ylab("Frequency")+  
            ggtitle("Top 5 Publications and number of posts they have written")+
            geom_text(stat = 'count',aes(label=..count..),hjust=0.5)

# plotting the top 5 publications and their claps
g2_Top5_p<-Top5_pub[,c("publication","claps")] %>%
            group_by(publication) %>%
            summarise_each(funs(sum)) %>%
            ggplot(.,aes(x=str_wrap(publication,15),y=claps))+
            geom_bar(stat="identity")+
            xlab("Publication")+ylab("Claps")+  
            ggtitle("Total number of Claps they got")+
            coord_flip()+geom_text(aes(label=claps),hjust =0.5 )

# plotting two plots at same grid
grid.arrange(g1_Top5_p,g2_Top5_p,ncol=1)
```

# Word Cloud for the Titles from Top 10 Authors

Word cloud from the titles of the posts by Top 10 authors 
of most number of posts is below. The words thing, read, data, drone and
new are with highest mentions. Where words such as big, telecom and tech
are next in the line with higher amount of posts. In restrictions I have 
considered that this word cloud will have 1500 words and only if a word
atleast holds the frequency of 10.

Well, I could clearly see that there cannot be 1500 words here.

```{r word cloud for Top 10 Authors,fig.width=6,fig.height=6}
#convert into data frame
Top10_author<-data.frame(Top10_author)

# Calculate Corpus
Top10_author.Corpus<-Corpus(VectorSource(Top10_author$title))

# clean the data
Top10_author.Clean<-tm_map(Top10_author.Corpus,PlainTextDocument)
Top10_author.Clean<-tm_map(Top10_author.Corpus,tolower)
Top10_author.Clean<-tm_map(Top10_author.Clean,removeNumbers)
Top10_author.Clean<-tm_map(Top10_author.Clean,removeWords,stopwords("english"))
Top10_author.Clean<-tm_map(Top10_author.Clean,removePunctuation)
Top10_author.Clean<-tm_map(Top10_author.Clean,stripWhitespace)
Top10_author.Clean<-tm_map(Top10_author.Clean,stemDocument)

# save as png
#png(filename = "wordcloud1.png",width = 1024,height = 768)
# plot the word cloud
wordcloud(Top10_author.Clean,max.words = 1500,min.freq = 10,
          colors = brewer.pal(11,"Spectral"),random.color = FALSE,
          random.order = TRUE)
```

# Word Cloud for the Titles from Top 5 publications

This word cloud also has similar restrictions for number of words and
minimum frequency for a word. Words such as data, learn, use, machin, network,
deep, scienc and artifici have most amount of frequency. Further, words such as 
neural, intellig, chatbot, part and python are also with significant
amount of frequency. Here we can see clearly see there can be more than 1000 words.

```{r word cloud for Top 5 publications,fig.width=6,fig.height=6}
#convert into data frame
Top5_pub<-data.frame(Top5_pub)

# Calculate Corpus
Top5_pub.Corpus<-Corpus(VectorSource(Top5_pub$title))

#clean the data
Top5_pub.Clean<-tm_map(Top5_pub.Corpus,PlainTextDocument)
Top5_pub.Clean<-tm_map(Top5_pub.Corpus,tolower)
Top5_pub.Clean<-tm_map(Top5_pub.Clean,removeNumbers)
Top5_pub.Clean<-tm_map(Top5_pub.Clean,removeWords,stopwords("english"))
Top5_pub.Clean<-tm_map(Top5_pub.Clean,removePunctuation)
Top5_pub.Clean<-tm_map(Top5_pub.Clean,stripWhitespace)
Top5_pub.Clean<-tm_map(Top5_pub.Clean,stemDocument)

# save as png
#png(filename = "wordcloud2.png",width = 1024,height = 768)
# plot the word cloud
wordcloud(Top5_pub.Clean,max.words = 1500,min.freq = 10,
          colors = brewer.pal(11,"Spectral"),random.color = FALSE,
          random.order = TRUE)
```


