---
title: "ThanksGiving week 34"
author: "M.amalan"
date: "November 20, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 10,fig.height = 10)
library(ggplot2)
library(scales)
library(ggthemr)
library(stringr)
library(gridExtra)

load("Thanksgiving_data.RData")
ggthemr("grape")
attach(Thanksgiving)

#subset the people who said yes for celebrating thanksgiving
Thanksgiving_Yes<-subset(Thanksgiving,celebrate=="Yes")
attach(Thanksgiving_Yes)

#subset the people who said no for celebrating thanksgiving
Thanksgiving_No<-subset(Thanksgiving,celebrate=="No")
attach(Thanksgiving_No)
```

Packages : ggplot2, ggthemr, stringr, scales, gridExtra
TidyTuesday : Week 34
Data : thanksgiving_meals

Saying Prayer and Regions for all: 

East South Central, South Atlantic, West South Central have respective percentages of 76.7%, 69.2% and 69.2% of 
saying prayers for Thanksgiving dinner. While East North Central, Middle Atlantic, Mountain and West North Central
all have percentages for saying prayer above 50 but less than 65. Finally, regions New England and Pacific have the 
highest percentages of not saying prayer for Thanksgiving with 51.7 and 41.8. In a special note, 52.5% have not 
answered which region they are from, and in this lot 25.4% would pray and 22% would not.
                            

```{r Prayer and Regions of all}
plot_all<-ggplot(Thanksgiving,aes(x=str_wrap(us_region,7),fill=prayer))+
          xlab("Regions")+ylab("Percentage")+ scale_y_continuous(labels = percent)+
          geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..]), position="stack" ) +
          geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..]) ),
                    stat="count", position=position_stack(1), vjust=1)+ggtitle("Thanksgiving and Praying with All",subtitle = "Some have not answered")

print(plot_all)

ggsave(plot_all,width = 10,height = 10,dpi = 300,filename = "Prayer_Regions_all.png")
```

Saying Prayer and Regions for people who celebrate Thanksgiving:

If we remove the people who do not celebrate Thanksgiving it is clear that the previous percentages have
changed. Further, East South Central, South Atlantic, West South Central percentages have increased to 
respectively 82.1, 72.9 and 74.1. While East North Central, Middle Atlantic, Mountain and West North Central
all have percentages for saying prayer above 60 but less than 70. Finally, regions New England and Pacific have the 
highest percentages of not saying prayer for Thanksgiving with 54.5 and 46.9. When we remove the people who do not
celebrate thanksgiving the people who have not answered which region they are from has decreased to 42.9%. while 
others two have increased.


```{r Prayer and Regions of people who celebrate Thanksgiving}
plot_yes<-ggplot(Thanksgiving_Yes,aes(x=str_wrap(us_region,7),fill=prayer))+
          xlab("Regions")+ylab("Percentage")+ scale_y_continuous(labels = percent)+
          geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..]), position="stack" ) +
          geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..]) ),
                    stat="count", position=position_stack(1), vjust=1)+ggtitle("Everyone who celebrates Thanksgiving but praying !!!",
                                                                               subtitle ="Still some of them do not wish to say where they are from" )

print(plot_yes)

ggsave(plot_yes,width = 10,height = 10,dpi = 300,filename = "Prayer_Regions_Yes.png")
```

People who do not celebrate Thanksgiving:

20.5% people who do not celebrate thanksgiving are from pacific, while lease percentage amount goes to
New England and West North Central with 3.8. Regions such as Middle Atlantic, Southern Atlantic show 
percentages respectively 17.9 and 14.1. Finally, Other areas show percentages above 5 and less than 8.
Considerably 12.8% of people have not answered which region they are from.

You can look at the counts see if they are with the percentages.

```{r People who do not celebrate Thanksgiving}
plot_No<-ggplot(Thanksgiving_No,aes(x=str_wrap(us_region,7)))+
          xlab("Regions")+ylab("Frequency")+ 
          geom_bar(aes(y = (..count..)/sum(..count..))) +
          geom_text(aes(y = ((..count..)/sum(..count..)), label = scales::percent((..count..)/sum(..count..))), stat = "count", vjust = -0.25) +
          scale_y_continuous(labels = percent) +ggtitle("Percentage perspective look at Regions of who do not celebrate Thanksgiving")

ggsave(plot_No,width = 10,height = 10,dpi = 300,filename = "Prayer_Regions_No.png")

plot_No_count<-ggplot(Thanksgiving_No,aes(x=str_wrap(us_region,7)))+
          xlab("Regions")+ylab("Frequency")+ 
          geom_bar(aes(y = (..count..))) +
          geom_text(aes(y = ((..count..)), label = (..count..)), stat = "count", vjust = -0.25)+
          ggtitle("Counts perspective look at Regions of who do not celebrate Thanksgiving")

grid.arrange(plot_No,plot_No_count,ncol=1)

ggsave(plot_No_count,width = 10,height = 10,dpi = 300,filename = "Prayer_Regions_No.png")
```




