---
title: "Week_1_2019_TidyTuesday"
author: "M.amalan"
date: "January 1, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 12,fig.height = 12,message = FALSE,
                      warning = FALSE)
```

```{r Prepare to TidyTuesday}
# load the necessary packages
library(tidyverse)
library(lubridate)
library(kableExtra)
library(ggthemr)

#load the ggthemr
ggthemr("flat dark")

# load the data set
tidytuesday_tweets<-readRDS("tidytuesday_tweets.rds")
```

# #Tidytuesday Tweets

Using plots and Tables to express the #TidyTuesday data-set. 
You can obtain the dataset from [here.](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-01-01)

## Earliest Tweet

The first tweet is on April 2nd and it has 156 favorites and 
64 retweets, where the tweet is from Thomas Mock and the next 3 tweets are also 
from him. 


```{r Earliest tweet}
#tidytuesday_tweets[order(tidytuesday_tweets$created_at),c(3,4,13,14,71)] %>%
#  head(5) %>%
#  kable()  %>%
#  kable_styling()
```

## Any Verified Profiles ?

There are only 3 verified profiles where Hadley Wickham has the highest
amount of followers of 76469, where that tweet has 61 favorites but no
retweets. Other profiles are Civis Analytics and grspur, but both of them
have friends above 600 counts, but Hadley Wickham friends close to 290.

```{r}
#subset(tidytuesday_tweets[,c(4,13,14,76,77,82)],verified==TRUE) %>%
#  kable() %>%
#  kable_styling()
```

## Source of Tweets

Close to 1050 tweets are done by the web client and other clients
such as Android and Iphone have tweet counts of respectively 106 and 233.
Other sources include oddly Instagram, Facebook, WordPress and LinkedIn, which
I am naming because of their popularity.

```{r Source of Tweets}
ggplot(tidytuesday_tweets,aes(fct_infreq(source)))+
  geom_bar()+coord_flip()+
  geom_text(stat = "count",aes(label=..count..),hjust=-0.25)+
  ylab("Frequency")+xlab("Types of Sources")+
  ggtitle("Source of Tweets")
```

## Tweets Per Month

Beginning of #TidyTuesday we have 293 tweets on the month of April. Even
though over the next months the number of tweets are decreasing this is not
the case in October. Lowest number of tweets are recorded in September with
115 tweets. 

```{r Tweets Per Month}
ggplot(tidytuesday_tweets,
       aes(x=month(tidytuesday_tweets$created_at)))+
  geom_bar()+
  geom_text(stat = "count",aes(label=..count..),vjust=-0.15)+
  scale_x_continuous(breaks = seq(1,12),labels = seq(1,12))+
  ylab("Frequency")+ xlab("Months")+
  ggtitle("Tweet Counts By Month")
```

## Most Tweets By Screen Name

There are 30 twitter users if we consider the accounts that have tweeted
more than or equal to 10 tweets under the hashtag "TidyTuesday".
Thomas Mock has tweeted most which is 172 including retweets, and 
the second place goes to R4DScommunity with 92 tweets. All the 
other users have individually less than 40 tweets.

```{r Most Tweets By screen name}
tidytuesday_tweets %>%
  group_by(screen_name) %>%
  filter(n() >= 10) %>%
ggplot(aes(x=fct_infreq(screen_name)))+
  geom_bar()+ coord_flip()+
  geom_text(stat = "count",aes(label=..count..),hjust=-0.15)+
  ylab("Frequency")+ xlab("Screen Name")+
  ggtitle("Screen Name with Most Tweets")

```

### Most Tweets By Screen Name and their Source

For the same plot if we consider the source for the tweets, 
it is clear that only seven sources were used. Mostly 
all of these users are using the web client, but some are using
the iPhone as well. R4DS community does more tweeting through iPhone
than TweetDeck. TweetDeck is a simple way of handling multiple twitter accounts
at the same time. Tidyyourworld account only uses Android and WeAreRLadies uses
only TweetDeck.

```{r Most Tweets By Screen name and their Source}
tidytuesday_tweets %>%
  group_by(screen_name) %>%
  filter(n() >= 10) %>%
ggplot(aes(x=fct_infreq(screen_name),fill=source))+
  geom_bar(position = "stack",stat="count")+ 
  coord_flip()+
  geom_text(stat = "count",aes(label=..count..),hjust=1,
            position = position_stack())+
  ylab("Frequency")+ xlab("Screen Name")+
  ggtitle("Screen Name with Most Tweets and their Source")
```

### Most Tweets By Screen Name with their Retweet Counts

Of the Top 30 users with most amount of tweets the highest
amount of retweets is to a tweet from WeAreRLadies and
it is 95. There are more outliers from Thomas Mock. and the 
highest range is to the user drob. Most from this top 30 users
have the range between 0 and 10.

```{r Most Tweets By Screen name with Retweets}
tidytuesday_tweets[,c("screen_name","retweet_count")] %>%
  group_by(screen_name) %>%
  filter(n() >= 10) %>%
ggplot(.,aes(x=fct_infreq(screen_name),y=retweet_count))+
  geom_boxplot()+ coord_flip()+
  scale_y_continuous(breaks = seq(0,100,5),labels = seq(0,100,5))+
  ylab("Retweets")+ xlab("Screen Name")+
  ggtitle("Screen Name with Most Tweets and their Retweets Count")

```

### Most Tweets By Screen Name with their Favorite Counts 

Similarly Thomas Mock has more outliers, and the highest range is
to the user drob. Second place for outliers goes to R4DScommunity user.
Close to 500 favorites are counted to a tweet by drob and second place is
to a tweet by WeAreRladies with favorite counts slightly above 450.

```{r Most Tweets By Screen name with Favorite counts}
tidytuesday_tweets[,c("screen_name","favorite_count")] %>%
  group_by(screen_name) %>%
  filter(n() >= 10) %>%
ggplot(.,aes(x=fct_infreq(screen_name),y=favorite_count))+
  geom_boxplot()+ coord_flip()+
   scale_y_continuous(breaks = seq(0,500,25),labels = seq(0,500,25))+
  ylab("Favourites Count")+ xlab("Screen Name")+
  ggtitle("Screen Name with Most Tweets and their Favourties Count")

```

## Relationship between Favorite Counts vs Retweet Counts ?

Very clear positive correlation. Y scale ranges from 0 to 500,
where x scale range is from 0 to 100 and most of the data points
are centered around the range of 0 to 12 retweets and 0 to 60 Favorite.
a Few data data points are out of the above mentioned range.

```{r Scatter plot between favourite count vs retweet count}
ggplot(tidytuesday_tweets, 
       aes(x=retweet_count,y=favorite_count))+
  geom_point()+geom_smooth()+
  scale_x_continuous(breaks =seq(0,100,2) ,labels =seq(0,100,2))+
  scale_y_continuous(breaks =seq(0,500,10),labels =seq(0,500,10))+
  xlab("Retweets")+ylab("Likes")+
  ggtitle("Retweets Versus Likes")
```

## Relationship between Followers Count vs Friends Count ?

Here I have considers only twitter profiles which has followers count less 
than 5000 with friends count also less than 5000. The reason is to explain the 
relationship more clearly. Clearly most of the twitter profiles are has followers
less than 2000 with Followers also less than 2000. Clearly there are some profiles
with Followers count above 1000 but friends count less than 1000. Even though there 
are few profiles with less than 1000 followers but more than 1000 Friends

```{r Scatter plot between Followers count vs Friends count}
ggplot(subset(tidytuesday_tweets,
       followers_count < 5000 & friends_count < 5000), 
       aes(x=followers_count,y=friends_count))+
  geom_point()+geom_smooth()+
  scale_x_continuous(breaks =seq(0,5000,250),labels =seq(0,5000,250))+
  scale_y_continuous(breaks =seq(0,5000,250),labels =seq(0,5000,250))+
  xlab("Followers Count")+ylab("Friends Count")+
  ggtitle("Followers Count Versus Friends Count")
```

# Conclusion

My conclusion of the above plots and tables in point form

* Using tidyverse as usual is fun. 

* The Box plots for several variables in the same plot is easy
  for the use of comparison.
  
* The Scatter plots are nice to understand the relationship among two 
  continuous variables.
  
* The geom_smooth function is also very useful in modelling the data.  

# Further Analysis

* We can focus on the text variable which could be used for a word cloud.

* Further we can try to understand the hashtags with favorites and retweets.